---
title: "Mapping With More Data"
author: "Zixuan Liu, Zijie Huang"
date: "11/11/2020"
output:
  pdf_document: 
    latex_engine: xelatex
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
pacman::p_load("knitr","maps","tmap","sp","rgeos","sf","tidyverse","kableExtra")
```
# 1. Overview

## 1.1 Introduction

This project focuses on digging into the dataset of Public Assistance Funded Projects Details for hurricane from 2009 to 2018. Our main objective is to produce US county-level maps that shows declared hurricane. For each map, details about public assistence projects such as the number of projects and the amount of federal share obligated are shown. We built a shiny app to show the map.

## 1.2. Dataset: Public Assistance Funded Projects Details

Source: [**Public Assistance Funded Projects Details**](https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1)

# 2. Data Cleaning

## 2.1 Clean Disaster Declaration Summaries Dataset

The data was downloaded from the FEMA website. There are total 23 variables in this dataset. For our aim of the project, we filtered down to include only the hurricanes and the dates of interest. The new dataset call DisasterSummariesCleaned. The head of five observations and importnat variables are listed below:  

```{r echo=FALSE}
data1 <- read.csv("DisasterSummariesCleaned.csv", header = TRUE)
data1 <- data1 %>%
            select(c(1,2,7,11,19,24))
kable(head(data1,n=5)) %>%
kable_styling(font_size=9.5)
```


## 2.2 Clean Public Assistance Funded Projects Details Dataset

The data was downloaded from the FEMA website. There are total 22 variables in this dataset. For our aim of the project, we filtered down to include only the hurricanes and the dates of interest, and we took 11 variables out to make a new dataset call hurricaneClean. The head of five observations and selected variables are listed below:  

```{r echo=FALSE}
data2 <- read.csv("hurricaneClean.csv", header = TRUE)
data2 <- data2 %>% 
          select(c(-1,-4,-5,-7))
kable(head(data2,n=5)) %>%
kable_styling(font_size=7.5)
```

# 3. Mapping

## 3.1 Data Modification

In order to map these hurricanes, dataset had to be merged together with state and county mapping data from the "maps" R package. Below is the result of merged state data with county fips name and their longitude and latitude.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Load data
disaster <- read.csv("DisasterSummariesCleaned.csv",header = TRUE)
detail <- read.csv("hurricaneClean.csv",header = TRUE)

#data processing
#Get name by county.fips
county_fips <- county.fips

#Get state name + county name
disaster <- disaster %>%
  mutate(fips = fipsStateCode * 1000 + fipsCountyCode)
disaster <- left_join(disaster,county_fips, by = "fips")
disaster <- disaster %>% 
  select(c(26,24,2,4,5,11)) %>%
  mutate(ID = polyname) %>%
  separate(polyname,c("state","county"),sep = ",")
disaster <- na.omit(disaster)

#select states 
StatesInt <- state.name
StatesInt <- str_to_lower(StatesInt,locale = "en")

#Get county map
county <- st_as_sf(maps::map("county",StatesInt,plot=F,fill=T))

#Get subdata by condition
disaster_sub <- disaster %>% 
                  filter(year==2010,
                        state=="texas",
                        disasterNumber==1931,
                        paProgramDeclared==1)


    #Get filtered data
    
    disaster_sub <- disaster %>% 
      filter(year==2010,
             state=="texas",
             disasterNumber==1931,
             paProgramDeclared==1)
    
    detail_sub <- detail %>%
      filter(disasterNumber==1931) %>%
      mutate(ID = str_c(state,county,sep = ","))
    
    detail_sub2 <- detail_sub %>%
      group_by(ID) %>%
      summarise(total = sum(totalObligated))
    
    detail_sub <- left_join(detail_sub,count(detail_sub,ID),by = "ID")
    detail_sub <- left_join(detail_sub,detail_sub2,by = "ID")
    
    detail_sub <- detail_sub %>% 
      select(c(13,14,15)) %>%
      rename("Project Number" = "n","Federal Share Obligated" = "total") %>%
      mutate(ID = str_to_lower(ID,locale="en"))
    
    #Choose by condition
    county_sub <- subset(county,grepl("texas",county$ID))
    
    #Combine
    county_sub <- left_join(county_sub,disaster_sub,by = "ID")
    
    county_sub <- left_join(county_sub,detail_sub,by = "ID")
    
    #Deal with PA
    county_sub[is.na(county_sub)] <- 0
    
    county_sub$`Designated Counties` <- ifelse(county_sub$paProgramDeclared==1,"Public Assistance","No Designation")
    
    #Get x,y coordinates
    county_sub <- cbind(county_sub, st_coordinates(st_centroid(county_sub)))
    
    #Get county name
    county_sub <- county_sub %>%
      separate(ID,c("state","county"),sep = ",")
    
    #Filter again
    county_sub <- county_sub %>%
                    filter(state=="texas")
    
a=c("state","county","Project.Number","Federal.Share.Obligated","Designated.Counties","X","Y")
county_sub1 <- county_sub[a]

kable(head(county_sub1,n=5)) %>%
kable_styling(font_size=6)
```

## 3.2 Basic Map

Once the data was cleaned, the next step was to map the data for each hurricane on separate state maps using the ggplot2 R package. An example of state Texas can be previewed below.

```{r echo=FALSE}
 ggplot() +
      geom_sf(data = county_sub) +
      scale_fill_manual(values = col) +
      ggtitle("Example of State Texas") +
      theme(axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            legend.title = element_text(size = 10),
            legend.text = element_text(size = 10),
            plot.title = element_text(hjust = 0.5, face = "bold", size = 10))
```


## 3.3 Feature Added

Next step is to add Public Assistance Project, Project Number and Federal Share Obligated feature to our map. Below is an example for state Taxes.

```{r echo=FALSE, fig.width = 20,fig.height = 20,fig.align  = 'center'}
#Set color depends on result
    col <- c("white","lightgoldenrod1")
    
    #Plot
    ##Creaet title
    b <- unique(county_sub$disasterNumber)
    b <- b[str_length(b)>2]
    c <- unique(county_sub$declarationType)
    c <- c[str_length(c)>1]
    d <- str_sub(unique(county_sub$declarationDate),1,10)
    d <- d[str_length(d)>2]
    e <- unique(county_sub$state)
    e <- e[str_length(e)>2]
    title <- str_c("FEMA-",b,"-",c,", ",str_to_title(e)," Disaster Declaration as of ",d,sep = "")
    
#Public Assistance Project
    ggplot() +
      geom_sf(data = county_sub,aes(fill = Designated.Counties)) +
      scale_fill_manual(values = col) +
      ggtitle(title) +
      geom_text(data = county_sub, aes(X, Y, label = county), size = 3.5, fontface = "bold") +
      theme(axis.title.x = element_blank(),
            axis.title.y = element_blank(),
            axis.text.x=element_blank(),
            axis.text.y=element_blank(),
            legend.title = element_text(size = 20),
            legend.text = element_text(size = 20),
            plot.title = element_text(hjust = 0.5, face = "bold", size = 40)
      )

#Project Number
    ggplot() +
        geom_sf(data = county_sub,aes(fill = Project.Number)) +
        scale_fill_gradient(low="white", high="red") +
        ggtitle(title) +
        geom_text(data = county_sub, aes(X, Y, label = county), size = 3.5, fontface = "bold") +
        theme(axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20),
              plot.title = element_text(hjust = 0.5, face = "bold", size = 40)
        )
      
#Federal Share Obligated     
    ggplot() +
        geom_sf(data = county_sub,aes(fill = Federal.Share.Obligated)) +
        scale_fill_gradient(low="white", high="red") +
        ggtitle(title) +
        geom_text(data = county_sub, aes(X, Y, label = county), size = 3.5, fontface = "bold") +
        theme(axis.title.x = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              legend.title = element_text(size = 20),
              legend.text = element_text(size = 20),
              plot.title = element_text(hjust = 0.5, face = "bold", size = 40)
        )
   
```

# 4. Shiny App

Finally, a shiny application was created in order to further explore the maps. The shiny app can be viewed [**here**](https://zhuangbc.shinyapps.io/FEMA_Project/)
 